# Functions \label(sec:functions)

Functions in R allow use a set of commands that we can keep calling repeatedly. The goal of a function is to fulfill a certain task, typically relying on inputs from a user. The rest of the problems in this short book has many functions that we ask you to write, so here, we start with some simple functions.

## Questions

1.  Write a function to print `Hello World` if user inputs 1 and `Mellow World` if user inputs 0

    ```{r}
    hello_or_mellow <- function(choice)
    {
      if(choice == 1)
      {
        print("Hello World")
      }
      if(choice == 0)
      {
        print("Mellow World ")
      }
    }
    ```

    \

2.  Write a function to return `Hello World` if user inputs 1 and `Mellow World` if user inputs 0. Think about the difference between this function and that previous one.

    ```{r}
    hello_or_mellow <- function(choice)
    {
      if(choice == 1)
      {
        return("Hello World")
      }
      if(choice == 0)
      {
        return("Mellow World ")
      }
    }
    ```

    \

3.  Consider a scenario where a garden is circular, and you need to calculate the area of the garden. Write an R function that takes the radius of the garden as input and returns the area. How would you modify this function if the garden is elliptical instead of circular?

    ```{r}
    # Function to calculate area of a circular garden
    calculate_circular_garden_area <- function(radius) 
    {
      area <- pi * radius^2
      return(area)
    }

    # Function to calculate area of an elliptical garden
    calculate_elliptical_garden_area <- function(radius1, radius2) {
      area <- pi * radius1 * radius2
      return(area)
    }
    ```

    \

4.  Imagine you have two sets of exam scores, one from the first midterm and the other from the second midterm. Write an R function that takes both sets of scores as inputs and returns the set with the higher average score.

    ```{r}
    compare_average_scores <- function(scores1, scores2) 
    {
      if (mean(scores1) > mean(scores2)) 
      {
        return(scores1)
      } else {
        return(scores2)
      }
    }
    ```

    \

5.  (Not a function question) Create two vectors `vec1` and `vec2`, where `vec1` contains 1 and [Your Name]{.underline} while `vec2` contains your JEE Advance/JAM Rank and Answer to the Question as `TRUE` or `FALSE`

    > Dootika Ma'am will give all $A$s in this course

    Can such `vec1` and`vec2` be constructed? If yes, what are the datatype of 1 in `vec1`, and Answer to the question in `vec2`? Use `typeof()` function for finding out.

    ```{r}
    # Define vec1 with 1 and my name
    vec1 <- c(1, "Problems")

    # Define vec2 with my JEE Advanced Rank (assuming rank as 10000) and the answer to the question 
    jee_rank <- 10000  # Replace this with your actual JEE Advanced Rank/ JAM Rank
    answer <- TRUE   

    vec2 <- c(jee_rank, answer)

    # Print vec1 and vec2 to verify
    vec1
    vec2

    # Check the datatype of 1 in vec1
    typeof(vec1[1])

    # Check the datatype of the boolean answer in vec2
    typeof(vec2[2])
    ```
